### 2019-07-07 paper/header.tex

% **************GENERATED FILE, DO NOT EDIT**************

\title{HydroPowerModels.jl: A Julia/JuMP Package for Hydrothermal Economic Dispatch Optimization}

\author[1, 2]{Andrew W. Rosemberg}
\author[1, 2]{Alexandre Street}
\author[1, 2]{Joaquim D. Garcia}
\author[1, 3]{Thuener Silva}
\author[1, 3]{\\ Davi M. Valladão}
\author[4]{Oscar Dowson}
\affil[1]{Laboratory of Applied Mathematical Programming and Statistics (LAMPS)}
\affil[2]{Department of Eletrical Engineering, PUC-RIO}
\affil[3]{Department of Industrial Engineering, PUC-RIO}
\affil[4]{Department of Industrial Engineering, Northwestern University, Evanston}

\keywords{Julia, Optimization, Energy, Stochastic}